The polynomial representation of the type $A_{n - 1}$ rational Cherednik algebra in characteristic $p \mid n$

### 2015-05-19 formalwriteup.tex
Clean up linear independence, make generic c explicit

Define the polynomials 
\[
g(z)=\prod_{j=0}^{n-1} (1-x_jz) \qquad \text{ and } \qquad F(z) =\left(\sum_{k=0}^{p-1} \binom{c}{k} (g(z)-1)^k\right).
\]
In terms of $g(z)$ and $F(z)$, we have $F_i(z) = \frac{F(z)}{1-x_iz}$ and $f_i = [z^p] \frac{F(z)}{1-x_iz}$.  We will show that $f_i$ are singular vectors.

@@

\begin{proof}
We see easily that 
\[
\frac{\partial g}{\partial z} = g(z) \sum_j \frac{-x_j}{1-x_jz}.
\]
We now consider $\frac{\partial F}{\partial z}$. We compute
\begin{align*}
\frac{\partial F}{\partial z}%&=\frac{\partial}{\partial z}\left(\sum_{k=0}^{p-1} \binom{c}{k} (g-1)^k\right)\\
&=\sum_{k=1}^{p-1}k\binom{c}{k}(g-1)^{k-1}\frac{\partial g}{\partial z}\\
%&=\left(\sum_j \frac{-x_j}{1-x_jz}\right)\left(\sum_{k=1}^{p-1}k\binom{c}{k}(g-1)^{k-1}\right)g(z)\\
%&=\left(\sum_j \frac{-x_j}{1-x_jz}\right)\left(\sum_{k=0}^{p-2}(k+1)\binom{c}{k+1}(g-1)^{k}\right)g(z)\\
&=-\sum_j \frac{x_j}{1-x_jz}\sum_{k=0}^{p-2}c\binom{c-1}{k}(g-1)^{k}(g-1+1)\\
%&=\left(\sum_j \frac{-x_j}{1-x_jz}\right)\left(\left(\sum_{k=0}^{p-2}c\binom{c-1}{k}(g-1)^{k}\right)+\left(\sum_{k=0}^{p-2}c\binom{c-1}{k}(g-1)^{k+1}\right)\right)\\
&=-\sum_j \frac{x_j}{1-x_jz}\left(\sum_{k=0}^{p-2}c\binom{c-1}{k}(g-1)^{k}+\sum_{k=1}^{p-1}c\binom{c-1}{k-1}(g-1)^{k}\right)\\
&=-\sum_j \frac{x_j}{1-x_jz}\left(-c\binom{c-1}{p-1}(g-1)^{p-1}+\sum_{k=0}^{p-1}c\binom{c}{k}(g-1)^{k}\right)\\
&=-\sum_j \frac{cx_j}{1-x_jz}F(z)+\sum_j \frac{x_j}{1-x_jz}c\binom{c-1}{p-1}(g-1)^{p-1}.
\end{align*}
Defining the formal power series
\[
V(z)=\sum_j \frac{x_j}{1-x_jz}c\binom{c-1}{p-1}(g-1)^{p-1},
\]
we see that 
\[
F'(z) = V(z) - \sum_{j=0}^{n-1} \frac{cx_j}{1-x_jz} F(z).
\]
It remains only to show that $[z^l]V(z)=0$ for $l=0,\dots,p-1$.  Consider the expansion 
\[
g(z)=\prod_{j=0}^{n-1} (1-x_jz) = 1-z\left(\sum_i x_i\right)+z^2e_2(x_0, \ldots, x_{n - 1})+\dots+z^ne_n(x_0, \ldots, x_{n - 1})
\]
for some polynomials $e_2, \ldots, e_n$.  Recalling that $\sum_i x_i=0$ by hypothesis, we see that $z^2 \mid g(z)-1$. By definition of $V(z)$, we see that $(g-1)^{p-1} \mid V(z)$, hence $z^{2p-2} \mid V(z)$. Since $p \ge 2$, this implies $z^p \mid V(z)$, completing the proof.
\end{proof}

@@

\begin{proposition}\label{prop:linind} The $f_i$ for $i=0,\dots,n-2$ are linearly independent degree $p$ homogeneous polynomials.
\end{proposition}

\begin{proof} 
We recall that $F_i(z)=\frac{1}{1-x_iz}\left(\sum_{k=0}^{p-1} \binom{c}{k} (g-1)^k\right)=\left(\sum_{k=0}^\infty x_i^kz^k\right)\left(\sum_{k=0}^{p-1} \binom{c}{k} (g-1)^k\right)$. It is then clear that $f_i$ is a homogeneous polynomial in the $x_j$ of degree $p$, since $[z^k]F_i(z)$ is a homogeneous polynomial in the $x_j$ of degree $k$ for all $k$ (this follows from the fact that this is true in both multiplicands in $F_i$).

Assume $\sum_{i=0}^{n-2} \lambda_if_i=0$ for some $\lambda_i \in k$. Then in particular this must be true when we set $x_j=1$ and $x_i=0$ for $i \ne j,i < n-1$. (We let $x_{n-1}=-1$ so that $\sum x_i=0$.) Note that if this is the case that $g(z)=(1-z)(1+z)=1-z^2$, so 
\[
F_j(z)=\left(\sum_{k=0}^\infty z^k\right)\left(\sum_{k=0}^{p-1} \binom{c}{k} (-z^2)^k\right)
\]
and 
\[
F_i(z)=\left(\sum_{k=0}^\infty 0^kz^k\right)\left(\sum_{k=0}^{p-1} \binom{c}{k} (-z^2)^k\right)=\sum_{k=0}^{p-1} \binom{c}{k} (-z^2)^k
\]
for $n-1>i \ne j$.

We first consider the case $p=2$. In this case $[z^2]F_j(z)=1-c$ and $[z^2]F_i(z)=-c$. By setting $x_j=1,x_{n-1}=-1$ and $x_i=0$ for $i \ne j, i \ne n-1$ for all $j=0,\dots,n-1$ in succession, we find that 
\[
\lambda_j-c\sum_{i=0}^{n-2} \lambda_i = 0 \text{ for all }j.
\]
Therefore for all $j$, $\lambda_j=c\sum_{i=0}^{n-2} \lambda_i$. Then in particular all of the $\lambda_i$ are equal to some $\lambda \in k$ with $(1-c(n-1))\lambda = 0$. Since $c$ is generic, we can assume $(1-c(n-1)) \ne 0$, so then $\lambda=0$. Therefore $\lambda_j=0$ for all $j$; then the $f_i$ for $i=0,\dots,n-2$ are linearly independent as desired.

We now suppose $p>2$. Then we have $[z^p]F_j(z)=f_j=\sum_{k=0}^{(p-1)/2} (-1)^k\binom{c}{k}=\binom{c-1}{(p-1)/2}$ and $[z^p]F_i(z)=f_i=0$ for $n-1>i \ne j$ (since $p > 2$). Since $c$ is generic, we can assume $\binom{c-1}{(p-1)/2} \ne 0$. Then since $\sum_{i=0}^{n-2} \lambda_if_i=0$, we see that $\lambda_j\binom{c-1}{(p-1)/2}=0$; then we must have $\lambda_j=0$.

By setting $x_j=1,x_{n-1}=-1$ and $x_i=0$ for $i \ne j, i \ne n-1$ for all $j=0,\dots,n-1$ in succession, we find that $\lambda_j=0$ for all $j$. Therefore the $f_i$ for $i=0,\dots,n-2$ are linearly independent as desired.
\end{proof}

@@

\begin{proposition}\label{prop:ci}
The quotient $A/I$ is a complete intersection. 
\end{proposition}
\begin{proof}
It suffices to show that if $x_0,\dots,x_{n-1} \in k$ satisfy $f_i(x_0,\dots,x_{n-1})=0$ for all $i$, then $x_0=\dots=x_{n-1}=0$.  Suppose that the $x_i$ take the distinct values $\{y_0,\dots,y_{r-1}\}$, where $y_i$ occurs with multiplicity $m_i>0$ so that $g(z)=(1-y_iz)^{m_i}$.  In these terms, we have that 
\[
[z^p]\left(\frac{1}{1-y_iz}F(z)\right)= 0
\]
for $i = 0, \ldots, r - 1$ and wish to show that $r=1$ and $y_0=0$.

>>>>

Define the polynomials 
\[
g(z)=\prod_{j=0}^{n-1} (1-x_jz) \qquad \text{ and } \qquad F(z) = \sum_{k=0}^{p-1} \binom{c}{k} (g(z)-1)^k.
\]
In terms of $g(z)$ and $F(z)$, we have $F_i(z) = \frac{F(z)}{1-x_iz}$ and $f_i = [z^p] \frac{F(z)}{1-x_iz}$.  We will show that $f_i$ are singular vectors.

@@

\begin{proof}
We see easily that 
\[
\frac{\partial g}{\partial z} = g(z) \sum_j \frac{-x_j}{1-x_jz}.
\]
We now consider $\frac{\partial F}{\partial z}$. We compute
\begin{align*}
\frac{\partial F}{\partial z}%&=\frac{\partial}{\partial z}\left(\sum_{k=0}^{p-1} \binom{c}{k} (g-1)^k\right)\\
&=\sum_{k=1}^{p-1}k\binom{c}{k}(g-1)^{k-1}\frac{\partial g}{\partial z}\\
%&=\left(\sum_j \frac{-x_j}{1-x_jz}\right)\left(\sum_{k=1}^{p-1}k\binom{c}{k}(g-1)^{k-1}\right)g(z)\\
%&=\left(\sum_j \frac{-x_j}{1-x_jz}\right)\left(\sum_{k=0}^{p-2}(k+1)\binom{c}{k+1}(g-1)^{k}\right)g(z)\\
&=-\sum_j \frac{x_j}{1-x_jz}\sum_{k=0}^{p-2}c\binom{c-1}{k}(g-1)^{k}(g-1+1)\\
%&=\left(\sum_j \frac{-x_j}{1-x_jz}\right)\left(\left(\sum_{k=0}^{p-2}c\binom{c-1}{k}(g-1)^{k}\right)+\left(\sum_{k=0}^{p-2}c\binom{c-1}{k}(g-1)^{k+1}\right)\right)\\
&=-\sum_j \frac{x_j}{1-x_jz}\left(\sum_{k=0}^{p-2}c\binom{c-1}{k}(g-1)^{k}+\sum_{k=1}^{p-1}c\binom{c-1}{k-1}(g-1)^{k}\right)\\
&=-\sum_j \frac{x_j}{1-x_jz}\left(-c\binom{c-1}{p-1}(g-1)^{p-1}+\sum_{k=0}^{p-1}c\binom{c}{k}(g-1)^{k}\right)\\
&=-\sum_j \frac{cx_j}{1-x_jz}F(z)+\sum_j \frac{x_j}{1-x_jz}c\binom{c-1}{p-1}(g-1)^{p-1}.
\end{align*}
Defining the formal power series
\[
V(z)=\sum_j \frac{x_j}{1-x_jz}c\binom{c-1}{p-1}(g-1)^{p-1},
\]
we see that 
\[
F'(z) = V(z) - \sum_{j=0}^{n-1} \frac{cx_j}{1-x_jz} F(z).
\]
It remains only to show that $[z^l]V(z)=0$ for $l=0,\dots,p-1$.  Consider the expansion 
\[
g(z)=\prod_{j=0}^{n-1} (1-x_jz) = 1-z\sum_i x_i+z^2e_2(x_0, \ldots, x_{n - 1})+\dots+z^ne_n(x_0, \ldots, x_{n - 1})
\]
for some polynomials $e_2, \ldots, e_n$.  Recalling that $\sum_i x_i=0$ by hypothesis, we see that $z^2 \mid g(z)-1$. By definition of $V(z)$, we see that $(g-1)^{p-1} \mid V(z)$, hence $z^{2p-2} \mid V(z)$. Since $p \ge 2$, this implies $z^p \mid V(z)$, completing the proof.
\end{proof}

@@

\subsection{Proof of linear independence of $f_i$}

\begin{proposition}\label{prop:linind} 
For generic $c$, the $f_i$ for $i=0,\dots,n-2$ are linearly independent degree $p$ homogeneous polynomials.
\end{proposition} 
\begin{proof} 
We have the expansion
\[
F_i(z)=\frac{1}{1-x_iz}\sum_{k=0}^{p-1} \binom{c}{k} (g-1)^k=\sum_{k=0}^\infty x_i^kz^k\sum_{k=0}^{p-1} \binom{c}{k} (g-1)^k.
\]
Because for any $l$ the coefficient of $z^l$ in each factor is a homogeneous polynomial of degree $l$, we see that $[z^p]F_i(z)$ is homogeneous of degree $p$.

For linear independence, suppose that $\sum_{i=0}^{n-2} \lambda_if_i=0$ for some $\lambda_i \in k$.  Substitute $x_{n-1} = -1$, $x_j=1$ and $x_i=0$ for $i \ne j,i < n-1$ so that $g(z)=(1-z)(1+z)=1-z^2$ and hence
\[
F_j(z)=\sum_{k=0}^\infty z^k\sum_{k=0}^{p-1} \binom{c}{k} (-z^2)^k
\]
and 
\[
F_i(z)=\sum_{k=0}^\infty 0^kz^k\sum_{k=0}^{p-1} \binom{c}{k} (-z^2)^k=\sum_{k=0}^{p-1} \binom{c}{k} (-z^2)^k \text{ for $i < n - 1$, $i \neq j$}.
\]
If $p = 2$, we see that $[z^2]F_j(z)=1-c$ and $[z^2]F_i(z)=-c$, so varying $j$ implies that
\[
\lambda_j = c\sum_{i=0}^{n-2} \lambda_i \text{ for all }j.
\]
In particular, all $\lambda_i$ have common value $\lambda \in k$ solving $(1-c(n-1))\lambda = 0$, which for generic $c$ implies that $\lambda = 0$, giving linear independence.

If $p > 2$, we have 
\[
[z^p]F_j(z)=f_j=\sum_{k=0}^{(p-1)/2} (-1)^k\binom{c}{k}=\binom{c-1}{(p-1)/2}
\]
and $[z^p]F_i(z)=f_i=0$ for $i < n - 1$ and $i \neq j$. For generic $c$, we have $\binom{c-1}{(p-1)/2} \ne 0$, meaning that 
\[
\sum_{i=0}^{n-2} \lambda_if_i=0 = \lambda_j\binom{c-1}{(p-1)/2}=0
\]
implies $\lambda_j=0$.  Varying $j$ implies that $\lambda_j=0$ for all $j$, again yielding linear independence.
\end{proof}

@@

\begin{proposition}\label{prop:ci}
For generic $c$, the quotient $A/I$ is a complete intersection. 
\end{proposition}
\begin{proof}
It suffices to show that if $x_0,\dots,x_{n-1} \in k$ satisfy $f_i(x_0,\dots,x_{n-1})=0$ for all $i$, then $x_0=\dots=x_{n-1}=0$.  Suppose that the $x_i$ take the distinct values $\{y_0,\dots,y_{r-1}\}$, where $y_i$ occurs with multiplicity $m_i>0$ so that $g(z)=(1-y_iz)^{m_i}$.  In these terms, we have that 
\[
[z^p]\left(\frac{1}{1-y_iz}F(z)\right)= 0
\]
for $i = 0, \ldots, r - 1$ and wish to show that $r=1$ and $y_0=0$.
